Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increases Repair Time?

### 2015-03-11 data.tex
Outline added for Forrest

The section presents the data that we use to conclude that phase delay is a not a major
contributor to the time required to resolve an issue.

The Software Engineering Institute (SEI) at
Carnegie Mellon University explores methods for software process improvement.
One product from that research is  
Watts Humphrey's  Team Software ProcessSM (TSPSM)~\cite{tsp00}. TSP is an extension of Humphrey's early
work on Personal Software
ProcessSM (PSPSM)~\cite{psp05}.

Since 2004, one  of use (Nichols) has been been mentoring 171 software development teams around the world
as they  deploy TSP within their organizations.  
The projects
were mostly small to medium, with a median duration of 46 days
and a maximum duration of 90 days. Median team size was 7
people, with a maximum of 40. Hence,  total development 
effort\footnote{
Effort =  duration*teamSize/workDaysPerYear.} for
these projects ranged from 1.3 years (median) to 14.3 years (max).
The majority of
the projects were either web portals or banking systems in the US, South Africa, and Mexico. 
There were also some  medical devices (US, France, Japan, and Germany),  a few from a commercial 
computer-aided design system (developed by a world wide distributed team), and less that  
a dozen United States
Department of Defence projects.
 An anonymized version of that data is available in the PROMISE repository\footnote{\carter{Need to add}} and,
 for confidentiality restrictions, we cannot offer 
further details
on those project.

@@

@@

\subsection{Project Phases}
This paper studies the impact of phase delay on the time required to resolve issues.
The phased used in this paper are shown in \fig{waterfall}. Note that, in that figure:
\bi 
\item
Several  phases have the same  sub-activities of {\em review} and {\em inspect}\footnote{\bill{plz distinguish review and inspect. does one usually happen first?}}
\item Testing is divided into several statges \footnote{\bill{need a one
line description of test vs qualtest vs inttest vs systemTest vs AcceptTest. I tries some words in the Key to \fig{sedata}. dont know if i got it right.}}
\ei 
Also shown in \fig{waterfall} (bottom left) is the standard definition of an agile process~\cite{boehmturner03}:
Given some
some backlog of tasks, wgeb teams complete their current tasks, they select the next task(s) to complete. That selection process may use a variety of criteria
to prioritize which  tasks are selected (for more details on that selection process, see~\cite{me09j,port08,boehmturner03}). Tasks are completed in ``sprints'' that can last hours,
days, but rarely not more than weeks. Each day meet for brief ``scrum'' sessions to assess (and possibly alter) their current progress on the goals of the sprint.  
Agile teams race to generate releases
(in the continuous release model, releases can be generated on a daily basis, or even faster).  
Experience gained from those releases informs the discussion in the daily scrums which, in turn,
can inform the team's decisions on how to select and implement the next set of tasks for next sprint.
In fact, that experience can result in changing some/all of the tasks in the current backlog.

>>>>

We examine the phase delay effect bin
171 software projects from around the world from 
\bill{2005 to 2013} that used the Team Software Process (TSP).

\subsection{Overview of the Team Software Process}
The Software Engineering Institute (SEI) at
Carnegie Mellon University explores methods for software process improvement.
One product from that research is  
Watts Humphrey's  Team Software ProcessSM (TSPSM)~\cite{tsp00}. TSP is an extension of Humphrey's early
work on Personal Software
ProcessSM (PSPSM)~\cite{psp05}.

\sei{One paragraph summary on the key points of TSP execution? The following text should also be reviewed and better integrated.}

@@

\subsection{Projects in the Sample}
Since 2004, one of the authors (Nichols) has been been mentoring 171 software development teams around the world as they deploy TSP within their organizations.  
The projects
were mostly small to medium, with a median duration of 46 days
and a maximum duration of 90 days. Median team size was 7
people, with a maximum of 40. Hence,  total development 
effort\footnote{
Effort =  duration*teamSize/workDaysPerYear.} for
these projects ranged from 1.3 years (median) to 14.3 years (max).
The majority of
the projects were either web portals or banking systems in the US, South Africa, and Mexico. 
There were also some  medical devices (US, France, Japan, and Germany),  a few from a commercial 
computer-aided design system (developed by a world wide distributed team), and less that  
a dozen United States
Department of Defence projects.
 An anonymized version of that data is available in the PROMISE repository\footnote{\carter{Need to add}} and,
 for confidentiality restrictions, we cannot offer 
further details
on those project.

\subsection{Data Collection Process}
\sei{
\bi
    \item What are artifacts uploaded to the SEI tool?
    \item What are the units contained in those artifacts? Are they the ``issues'' as defined in the following paragraph?
    \item Is participation in the SEI data collection mandatory? i.e., is the data provided self selected?
    \item When is the data collected?
    \item What are screening/cleaning procedures, if any?
\ei
}

@@

\subsection{Project Phases}
This paper studies the impact of phase delay on the time required to resolve issues.
The phased used in this paper are shown in \fig{waterfall}. Note that, in that figure:
\bi 
\item
Several  phases have the same  sub-activities of {\em review} and {\em inspect}\footnote{\bill{plz distinguish review and inspect. does one usually happen first?}}
\item Testing is divided into several statges \footnote{\bill{need a one
line description of test vs qualtest vs inttest vs systemTest vs AcceptTest. I tries some words in the Key to \fig{sedata}. dont know if i got it right.}}
\ei

\todo{I recommend deleting the rest of this section. Replace with better descriptions of what the phases are.}
Also shown in \fig{waterfall} (bottom left) is the standard definition of an agile process~\cite{boehmturner03}:
Given some
some backlog of tasks, wgeb teams complete their current tasks, they select the next task(s) to complete. That selection process may use a variety of criteria
to prioritize which  tasks are selected (for more details on that selection process, see~\cite{me09j,port08,boehmturner03}). Tasks are completed in ``sprints'' that can last hours,
days, but rarely not more than weeks. Each day meet for brief ``scrum'' sessions to assess (and possibly alter) their current progress on the goals of the sprint.  
Agile teams race to generate releases
(in the continuous release model, releases can be generated on a daily basis, or even faster).  
Experience gained from those releases informs the discussion in the daily scrums which, in turn,
can inform the team's decisions on how to select and implement the next set of tasks for next sprint.
In fact, that experience can result in changing some/all of the tasks in the current backlog.

### 2015-03-11 lucas.tex
Outline added for Forrest

As part of a related research, we conducted a survey of software engineering practitioners and researchers on their views of commonly-head software engineering ''laws''. One of the laws questioned was drawn from from Glass~\cite{glass02} who cites Boehm \lucas{ref needed} as the source: ''Requirements errors are the most expensive to fix when found during production but the cheapest to fix early in development'' (RqtsErr for short). The survey was conducted in two phases using Amazon's Mechanical Turk. The first phase was conducted only with professional software engineers, and the second survey was conducted only with Program Committee (PC) members of the ESEC/FSE and ICSE conferences.

The respondents answered the following questions for the laws they were presented: 
\textbf{Agreement:} ''Based on your experience, do you agree that the statement above is correct?'' A Likert scale captured the agreement score from Strongly Disagree to Strongly Agree. A text box was provided to explain the answer.
\textbf{Applicability:} ''To the extent that you believe it, how widely do you think it applies among software development contexts?'' The possible answers were presented as a scale from -1 to 5:
\begin{itemize}
\item I don't know (-1)
\item This law does not apply at all (0)
\item Applies in a very narrow range of projects  (1)
\item Rarely applies (2)
\item Occasionally applies (3)
\item Very frequently applies (4)
\item Always applies (5)
\end{itemize}
Respondents were required to explain the applicability score in a text box.

In the ''law evaluation'' portion of the survey, participants were presented with the RqtsErr law and others drawn from \cite{glass02} and \cite{endres03}. The PC member survey contained an additional question: ''Do you agree with the following? 'In general, the longer errors are in the system (requirements errors, design errors, coding errors, etc.), the more expensive they are to fix''' (PhaseDelay for short). Responses were recorded using the \textbf{Agreement} question Likert scale.

Summary statistics for the agreement and applicability scores for the RqtsErr and PhaseDelay laws are presented in Figure~\ref{fig:survey_results}. Responses whose \textbf{applicability} response was ''I don't know'' are ommitted from analysis.

@@

While the practitioners strongly believed in the law, researchers were less firm. In reading the free response texts, most participants agreed that, unless the system and the affected requirement are simple/trivial, addressing errors may mean system redesign. Further, any changes to the production system have additional complexity (compared to changes while system in development) due to users, data existence, and architecture dependencies. The researchers who disagreed with the law generally asserted that requirements change can be expensive, but that depends on the process used (e.g., agile vs. waterfall) and the adaptability of the system architecture.

>>>>

We conducted a survey of software engineering practitioners and researchers on their views of commonly-held software engineering ``laws''. One of the laws questioned relates directly to phase delay: ``Requirements errors are the most expensive to fix when found during production but the cheapest to fix early in development'' (from Glass~\cite{glass02} who cites Boehm \lucas{need to find which exact reference he uses}). We abbreviate this law as RqtsErr. The survey was conducted in two phases using Amazon's Mechanical Turk. The first phase was conducted only with professional software engineers, and the second survey was conducted only with Program Committee (PC) members of the ESEC/FSE and ICSE conferences.

The respondents answered the following questions for the laws they were presented: \newline
\textbf{Agreement:} ``Based on your experience, do you agree that the statement above is correct?'' A Likert scale captured the agreement score from Strongly Disagree to Strongly Agree. A text box was provided to explain the answer. \newline
\textbf{Applicability:} ``To the extent that you believe it, how widely do you think it applies among software development contexts?'' The possible answers were presented as a scale from -1 to 5:
\begin{itemize}
\item I don't know (-1)
\item This law does not apply at all (0)
\item Applies in a very narrow range of projects  (1)
\item Rarely applies (2)
\item Occasionally applies (3)
\item Very frequently applies (4)
\item Always applies (5)
\end{itemize}
Respondents were required to explain the applicability score in a text box.

Participants were presented with the RqtsErr law and others drawn from \cite{glass02} and \cite{endres03}. The PC member survey contained an additional question: ``Do you agree with the following? 'In general, the longer errors are in the system (requirements errors, design errors, coding errors, etc.), the more expensive they are to fix''' (PhaseDelay for short). Responses were recorded using the \textbf{Agreement} question Likert scale.

Summary statistics for the agreement and applicability scores for the RqtsErr and PhaseDelay laws are presented in Figure~\ref{fig:survey_results} \lucas{To add results from other laws}. Responses whose \textbf{applicability} response was ''I don't know'' are ommitted from analysis.

@@

While the practitioners strongly believed in the law, researchers were less firm. In the free response texts, most participants agreed that addressing errors may mean system redesign unless the system and the affected requirement are simple/trivial. Further, any changes to the production system have additional complexity (compared to changes while system in development) due to users, data existence, and architecture dependencies. The researchers who disagreed with the law generally asserted that requirements change can be expensive, but that depends on the process used (e.g., agile vs. waterfall) and the adaptability of the system architecture.
